Presentation Title

### 2017-08-19 thesis/chapters/ch3-model.tex
Edit ch4 and 5

Our consensus protocol runs on top of Extended TrustChain.
It uses an unmodified asynchronous common subset (ACS) algorithm as the key building block.
The objectives of the protocol are to
    allow honest nodes always make progress (in the form of creating new CP blocks),
    compute correct consensus result in every round
    and have an unbiased election of facilitators.
Concretely, we define the following.
\begin{definition}
\label{def:consensus}
\textbf{\emph{TrustChain consensus}}

@@

\subsection{Background on asynchronous subset consensus}
\label{sec:acs-background}

@@

ACS is an especially useful primitive for blockchain systems.
It allows any party to propose a value and the result is the set union of all the proposed values by the majority.
Concretely, ACS needs to satisfy the following properties (adapted from~\cite{miller2016honey}).
\begin{definition}
\label{def:acs}
\textbf{\emph{Asynchronous subset consensus}}

@@

The TX protocol, shown in \Cref{alg:tx-proto}, is run by all nodes.
It is also known as True Halves, first described informally by Veldhuisen~\cite[Chapter~3.2]{truehalves}.
Nodes that wish to initiate a transaction calls $\textsf{new\_tx}(pk_v, m, txid)$ (\Cref{alg:new-tx}) with the intended counterparty $v$ identified by $pk_v$ and message $m$.
$txid$ should be a uniformly distributed random value, i.e. $txid \in_R \{0, 1\}^{256}$.
Then the initiator sends $\langle \texttt{tx\_req}, t_{u, h}\rangle$ to $v$.

@@

\begin{algorithm}
    \caption{The TX protocol which runs in the context of node $u$.}
    \label{alg:tx-proto}

@@

\subsection{Validation protocol}
Our validation protocol,
shown in~\Cref{alg:vd-proto},
is designed to classify transactions according to the aforementioned validity definition.
If $u$ wishes to validate some TX with ID $txid$ and counterparty $v$, it sends $\langle \texttt{vd\_req}, txid \rangle$ to $v$.
The desired properties of the validation protocol are as follows.
\begin{definition}
\label{def:validation}
\textbf{\emph{Properties of the validation protocol}}

\begin{enumerate}
    \item \emph{Agreement}:
        If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction,
        then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
    \item \emph{Validity}:
        The validation protocol outputs the correct result
        according to the aforementioned validity definition.
    % \item \emph{Unforgeability}:
    %     If some transaction is valid, it cannot be forged into an invalid transaction.
    %     If some transaction is invalid, it cannot be forged into a valid transaction.
    \item \emph{Liveness}:
        Any valid (invalid) transaction is marked as valid (invalid) eventually.
\end{enumerate}
\end{definition}

\begin{algorithm}
\caption{Validation protocol which runs in the context of $u$}
\label{alg:vd-proto}

>>>>

Our consensus protocol $\Pi_c$ runs on top of Extended TrustChain.
It uses an unmodified asynchronous common subset (ACS) algorithm as the key building block.
The objectives of the protocol are to
    allow honest nodes always make progress (in the form of creating new CP blocks),
    compute correct consensus result in every round
    and have an unbiased election of facilitators.
Concretely, we define the following.
\begin{definition}
\label{def:consensus}
\textbf{\emph{TrustChain consensus protocol}}

@@

\subsection{Background on asynchronous common subset}
\label{sec:acs-background}

@@

ACS is an especially useful primitive for blockchain systems.
It allows any party to propose a value and the result is the set union of all the proposed values by the majority.
Concretely, ACS needs to satisfy the following properties (adapted from~\cite{miller2016honey}).
\begin{definition}
\label{def:acs}
\textbf{\emph{Asynchronous common subset}}

@@

The TX protocol $\Pi_t$, shown in \Cref{alg:tx-proto}, is run by all nodes.
It is also known as True Halves, first described informally by Veldhuisen~\cite[Chapter~3.2]{truehalves}.
Nodes that wish to initiate a transaction calls $\textsf{new\_tx}(pk_v, m, txid)$ (\Cref{alg:new-tx}) with the intended counterparty $v$ identified by $pk_v$ and message $m$.
$txid$ should be a uniformly distributed random value, i.e. $txid \in_R \{0, 1\}^{256}$.
Then the initiator sends $\langle \texttt{tx\_req}, t_{u, h}\rangle$ to $v$.

@@

\begin{algorithm}
    \caption{The TX protocol $\Pi_t$ runs in the context of node $u$.}
    \label{alg:tx-proto}

@@

\subsection{Validation protocol}
Our validation protocol $\Pi_v$,
shown in~\Cref{alg:vd-proto},
is designed to classify transactions according to the aforementioned validity definition.
If $u$ wishes to validate some TX with ID $txid$ and counterparty $v$, it sends $\langle \texttt{vd\_req}, txid \rangle$ to $v$.
The desired properties of the validation protocol are as follows.
\begin{definition}
\label{def:validation}
\textbf{\emph{TrustChain validation protocol}}

A TrustChain validation protocol is correct if the following properties hold.
\begin{enumerate}
    \item \emph{Agreement}:
        If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction,
        then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
    \item \emph{Validity}:
        The validation protocol outputs the correct result
        according to the aforementioned validity definition.
    % \item \emph{Unforgeability}:
    %     If some transaction is valid, it cannot be forged into an invalid transaction.
    %     If some transaction is invalid, it cannot be forged into a valid transaction.
    \item \emph{Liveness}:
        Any valid (invalid) transaction is marked as valid (invalid) eventually.
\end{enumerate}
\end{definition}

\begin{algorithm}
\caption{Validation protocol $\Pi_v$ which runs in the context of $u$}
\label{alg:vd-proto}

### 2017-08-19 thesis/chapters/ch4-analysis.tex
Style improvements in thesis
Edit ch4 and 5

Up to this point, we described our system specification in detail.
Of course, specification alone does not establish any truths.
In this chapter, we evaluate our system analytically.
First, we show it has the desired properties.
That is, the properties of the consensus protocol and the validation protocol (\Cref{def:consensus} and \Cref{def:validation} respectively) should hold.
Then we analyse the performance, especially the throughput,
and show that it out performs early blockchain systems.
Finally, we consider the case where the $n \ge 3t + 1$ assumption is violated and show the effect of it in our system.

\section{Correctness in the presence of faults}
Our first objective is to show that \Cref{def:consensus} holds for our consensus protocol.
Then, building on top of it, we show \Cref{def:validation} holds for the validation protocol.
The resulting theorem shows that only using CP blocks in the consensus algorithm implies consensus on TX blocks.

@@

We begin our analysis by establishing truths on the four properties in \Cref{def:consensus},
namely agreement, validity, fairness and termination for an arbitrary round.
Using these results, we use mathematical induction to show that they hold for all rounds.

\begin{lemma}
\label{lemma:agreement}
For an arbitrary round $r$
if $\F_{r}$ is known by all correct nodes and one correct node outputs a list of facilitators $\F_{r+1}$,
then all correct nodes output $\F_{r+1}$.
\end{lemma}
\begin{proof}
The argument follows from the protocol description.
Given that $\F_r$ is known,
correct nodes will send CP blocks to all members in $\F_r$.
The ACS algorithm starts independently whenever the facilitator has $N - t$ valid CP blocks
(recall from \Cref{sec:consensus-phase} that invalid blocks are ones with an invalid signature or has a duplicate signature).
It cannot make progress until $n-t$ honest facilitators start algorithm,
but this eventually happens because there are $N - t$ correct nodes and all correct facilitator eventually receives $N -t$ valid CP blocks.
At the end of ACS, $\C_{r+1}$ is created and then broadcasted along with the signature of the facilitators.
Due to the agreement property of ACS (\Cref{def:acs}),
every correct node should receive at least $n - t$ valid signatures on the agreed $\C_{r+1}$.
Thus they use $\C_{r+1}$ to generate a new CP block and compute new facilitators.
Since $\textsf{get\_facilitators}(\cdot)$ is a deterministic algorithm and the input $\C_{r+1}$ is in agreement, the output $\F_{r+1}$ is also in agreement.
\end{proof}

\begin{lemma}
\label{lemma:validity}
For an arbitrary round $r$,
if $\F_r$ is known by all correct nodes and any correct node outputs $\F_{r+1}$,
then (a) $|\C_{r+1}| \ge N - t$ must hold for the $\C_{r+1}$ which was used to create $\F_{r+1}$,
(b) $\F_{r+1}$ must contain at least $n - t$ honest nodes and
(c) $|\F_{r+1}| = n$.
\end{lemma}
\begin{proof}
The validity follows from the validity property of ACS and the definition of our model,
namely $N \ge n + t$ and $n \ge 3t + 1$.
Given $\F_r$, since $N \ge n + t$ , there is at least $n$ nodes that would send their CP block to $\F_r$.
From the validity property of ACS, we know the output must contain the input of at least $n - 2t$ nodes.
But $n -t$ facilitators must have received $N - t$ valid CP blocks, so $|\C_{r+1}| \ge N-t$, this proves (a).
There are at least $n-t$ honest nodes in $\F_{r+1}$ which follows from our assumption, this proves (b).
Finally, since $N-t \ge (n+t) -t = n$ and $\textsf{get\_facilitators}(\cdot)$ outputs $n$ items, $|\F_{r+1}| = n$ and this proves (c).
\end{proof}

\begin{lemma}
\label{lemma:fairness}
For an arbitrary round $r$,
if $\F_r$ is known by all correct nodes then
every node with a CP block in $\C_{r+1}$ should have an equal probability to be elected as a facilitator in $\F_{r+1}$.
\end{lemma}
\begin{proof}
We already established that $|\C_{r+1}| \ge N - t \ge n$ from \Cref{lemma:validity}.
Then the proof directly follows from the random oracle model.
Recall that the luck value is computed using $\textsf{H}(\C_{r+1}, || pk_u)$.
Since $pk_u$ is unique for every node that has a CP block in $\C_{r+1}$,
the output of $\textsf{H}(\cdot)$ is uniformly random.
This effectively generates a random permutation
so every node has the same probability of being in the top $n$ for the ordered sequence,
namely the output of $\textsf{get\_facilitators}(\cdot)$.
\end{proof}

\begin{lemma}
\label{lemma:termination}
For an arbitrary round $r$,
if $\F_r$ is known by all correct nodes then
every correct node eventually outputs some $\F_{r + 1}$.
\end{lemma}
\begin{proof}
This follows directly from the properties of the channel (eventual delivery)
and the termination property of ACS.
That is, $\F_r$ eventually receives all the CP blocks required to begin ACS,
and then ACS eventually terminates.
Finally, the results are eventually disseminated to all the nodes.
\end{proof}

From Lemmas \ref{lemma:agreement}, \ref{lemma:validity}, \ref{lemma:fairness} and \ref{lemma:termination},
we have shown that the 4 properties of \Cref{def:consensus} holds when assuming the existence of some $\F_r$.
Thus, to proof the whole of \Cref{def:consensus},
we need to proof these 4 properties under the universal quantifier.
We do this using mathematical induction.

\begin{theorem}
\label{theorem:consensus}
For all rounds,
the consensus protocol satisfies agreement, validity, fairness and termination (\Cref{def:consensus}).
\end{theorem}
\begin{proof}
We proof using mathematical induction.

@@

For the inductive step,
we assume that the 4 properties hold in round $r$ and prove that they also hold in round $r + 1$.
Using Lemmas \ref{lemma:agreement}, \ref{lemma:validity}, \ref{lemma:fairness} and \ref{lemma:termination},
it directly follows from modus ponens that these properties hold for $r + 1$.
Due to the principals of mathematical induction, these properties hold for all $r$.
\end{proof}

\subsection{Correctness of the validation protocol}
The consensus protocol,
which ensures consensus on CP blocks and facilitators,
is the backbone for consensus on transactions.
In this section, we build on top of \Cref{theorem:consensus} to show that the properties except for liveness in \Cref{def:validation} can be satisfied.

\begin{theorem}
\label{theorem:validation-agreement}
% (Agreement of the validation protocol)
The validation protocol satisfies agreement and validity properties.
\end{theorem}
\begin{proof}
We proof the agreement property by contradiction.
From~\Cref{alg:get-validity},
we use the fact that $F_{u, i}$ and $t_{u, i}$ are valid.
Without loss of generality, suppose for some transaction with TX block $t$,
node $u$ decides \emph{valid} but node $v$ decides \emph{invalid}.
Then there exists a fragment $F' = \{ \dots, t', c'\}$ which $u$ received that contains a valid pair of $t$---$t'$.
There also exists a fragment $F'' = \{ \dots, t'', c''\}$ which $v$ received that does not contain or contains an invalid pair---$t''$.
In both cases, the $\textsf{get\_validity}(\cdot)$ function must have reached \Cref{line:valid-fragment}.
Due to \Cref{theorem:consensus}, we have $c' = c''$, otherwise the result would be \emph{unknown}.
Since $c' (= c'') = \langle \textsf{H}(t'), \dots \rangle$ we must have $\textsf{H}(t') = \textsf{H}(t'')$ and $t' \ne t''$ (because $t''$ is invalid).
In other words, the sender of $F''$ must be able to create some $t''$ that has the same digest as $t'$.
But this is only possible if the adversary can compute the inverse of $H(\cdot)$ with non-negligible probability.
Thus we have a contradiction and this completes the agreement proof.

@@

\subsection{Impossibility of liveness}
\Cref{theorem:validation-agreement} is a major result that allows significantly improved performance over traditional blockchain systems,
but it does not have all the properties of a typical Byzantine consensus protocol.
Now we show a negative result, where the liveness property of \Cref{def:validation} cannot be attained.
Meaning that transactions with adversaries cannot always be validated.

@@

\section{Performance analysis}
\label{sec:performance-analysis}
This section aims to analytically answer the scalability aspect of our research question.
That is, does the global throughput increase linearly with respect to the population size?
We begin by looking at the communication and time complexity of the consensus protocol,
and then the bandwidth requirement for a single transaction.
We build on top of those results to analyse the global throughput.

\subsection{Communication complexity of the consensus protocol}
\label{sec:cons-complexity}
The consensus protocol can be seen as three parts,
so the communication complexity is the sum of these parts.
The first part is when every node sends their CP block to all the facilitators, which is $O(Nn)$ since there are $N$ nodes and $n$ facilitators.
Or simply $O(N)$ if we consider $n$ as a constant.

@@

\section{Effect of a highly adversarial environment}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=1.0\textwidth]{hypergeometric}
  \caption{Plot of the probability of selecting more than $\lfloor \frac{n-1}{3} \rfloor$ black balls for
  different value of $t$ with $N$ fixed at 2000.}
  \label{fig:hypergeometric}
\end{figure}

@@

The bound is not tight, but it is useful for picking parameters.
For some fixed $n$, since $0 \le \alpha \le \lfloor \frac{n-1}{3} \rfloor/n < (\lfloor  \frac{n - 1}{3} \rfloor + 1) / n$,
the probability is maximum when the squared term is minimum at $\alpha = \lfloor \frac{n-1}{3} \rfloor/n$.
The probability is minimum when the squared term is maximum at $\alpha = 0$.
Hence, if $n$ is known, then we can pick a $\alpha$ such that the probability becomes small.
On the other hand, if $\alpha$ is fixed, but small, we may increase $n$ to achieve the same.
To put this into perspective,
suppose $n = 1000$ and $\alpha = 1/5$, i.e. a fifth of the nodes are malicious.
Then the probability to draw more black balls than the threshold is only $2.6 \times 10^{-16}$.

>>>>

In this chapter,
we evaluate our system analytically to check whether it correctly implements the TrustChain consensus protocol (\Cref{def:consensus})
and the TrustChain validation protocol (\Cref{def:validation}) in~\Cref{sec:correctness}.
Further, we analyse the performance and especially the throughput to understand the scalability characteristics in~\Cref{sec:performance-analysis}.
Finally, we consider the case where the $n \ge 3t + 1$ assumption is violated and show the effect of it in our system in~\Cref{sec:highly-adversarial}.

\section{Correctness in the presence of faults}
\label{sec:correctness}
Our first objective is to show that our consensus protocol $\Pi_c$ described in~\Cref{sec:cons-protocol} is in fact a TrustChain consensus protocol defined in~\Cref{def:consensus}.
Then, building on top of it, we show that our validation protocol $\Pi_v$ described in~\Cref{sec:vd-protocol} conforms to the TrustChain validation protocol defined in~\Cref{def:validation}.
The resulting theorem shows that only using CP blocks in the consensus algorithm implies consensus on TX blocks.

@@

We begin our analysis by establishing the four properties in \Cref{def:consensus},
namely agreement, validity, fairness and termination for an arbitrary round.
Using these results, we use mathematical induction to show that they hold for all rounds.

\begin{lemma}
\label{lemma:agreement}
If for an arbitrary round $r$,
$\F_{r}$ is known by all correct nodes and one correct node outputs a list of facilitators $\F_{r+1}$,
then all correct nodes output $\F_{r+1}$.
\end{lemma}
\begin{proof}
The argument follows from the protocol description.
Given that $\F_r$ is known,
correct nodes will send CP blocks to all members in $\F_r$.
The ACS algorithm starts independently whenever the facilitator has $N - t$ valid CP blocks
(recall from \Cref{sec:consensus-phase} that invalid blocks are ones with an invalid signature or has a duplicate signature).
It cannot make progress until $n-t$ honest facilitators start algorithm,
but this eventually happens because there are $N - t$ correct nodes and all correct facilitator eventually receives $N -t$ valid CP blocks.
At the end of ACS, $\C_{r+1}$ is created and then broadcasted along with the signature of the facilitators.
Due to the agreement property of ACS (\Cref{def:acs}),
every correct node should receive at least $n - t$ valid signatures on the agreed $\C_{r+1}$.
Thus they use $\C_{r+1}$ to generate a new CP block and compute new facilitators.
Since $\textsf{get\_facilitators}(\cdot)$ is a deterministic algorithm and the input $\C_{r+1}$ is in agreement, the output $\F_{r+1}$ is also in agreement.
\end{proof}

\begin{lemma}
\label{lemma:validity}
If for an arbitrary round $r$,
$\F_r$ is known by all correct nodes and any correct node outputs $\F_{r+1}$,
then (a) $|\C_{r+1}| \ge N - t$ must hold for the $\C_{r+1}$ which was used to create $\F_{r+1}$ and (b) $|\F_{r+1}| = n$.
\end{lemma}
\begin{proof}
The validity follows from the validity property of ACS and the definition of our model,
namely $N \ge n + t$ and $n \ge 3t + 1$.
Given $\F_r$, since $N \ge n + t$ , there is at least $n$ nodes that would send their CP block to $\F_r$.
From the validity property of ACS, we know the output must contain the input of at least $n - 2t$ nodes.
But $n -t$ facilitators must have received $N - t$ valid CP blocks, so $|\C_{r+1}| \ge N-t$ and this proves (a).
Since $N-t \ge (n+t) -t = n$ and $\textsf{get\_facilitators}(\cdot)$ outputs $n$ items, so $|\F_{r+1}| = n$ and this proves (b).
\end{proof}

\begin{lemma}
\label{lemma:fairness}
If for an arbitrary round $r$,
$\F_r$ is known by all correct nodes then
every node with a CP block in $\C_{r+1}$ should have an equal probability to be elected as a facilitator in $\F_{r+1}$.
\end{lemma}
\begin{proof}
We already established that $|\C_{r+1}| \ge N - t \ge n$ from \Cref{lemma:validity}.
Then the proof directly follows from the random oracle model.
Recall that the luck value is computed using $\textsf{H}(\C_{r+1}, || pk_u)$.
Since $pk_u$ is unique for every node that has a CP block in $\C_{r+1}$,
the output of $\textsf{H}(\cdot)$ is uniformly random.
This effectively generates a random permutation
so every node has the same probability of being in the top $n$ for the ordered sequence,
namely the output of $\textsf{get\_facilitators}(\cdot)$.
\end{proof}

\begin{lemma}
\label{lemma:termination}
If for an arbitrary round $r$,
$\F_r$ is known by all correct nodes then
every correct node eventually outputs some $\F_{r + 1}$.
\end{lemma}
\begin{proof}
This follows directly from the properties of the channel (eventual delivery)
and the termination property of ACS.
That is, $\F_r$ eventually receives all the CP blocks required to begin ACS,
and then ACS eventually terminates.
Finally, the results are eventually disseminated to all the nodes.
\end{proof}

From Lemmas \ref{lemma:agreement}, \ref{lemma:validity}, \ref{lemma:fairness} and \ref{lemma:termination},
we have shown that the four properties of \Cref{def:consensus} holds when assuming the existence of some $\F_r$.
Now we need to proof these properties under the universal quantifier on $r$.
We do this using mathematical induction.

\begin{theorem}
\label{theorem:consensus}
For all rounds, $\Pi_c$ implements a TrustChain consensus protocol (\Cref{def:consensus}).
\end{theorem}
\begin{proof}
We prove using mathematical induction.

@@

For the inductive step,
we assume that the four properties hold in round $r$ and prove that they also hold in round $r + 1$.
Using Lemmas \ref{lemma:agreement}, \ref{lemma:validity}, \ref{lemma:fairness} and \ref{lemma:termination},
it directly follows from modus ponens that these properties hold for $r + 1$.
Due to the principals of mathematical induction, these properties hold for all $r$.
\end{proof}

\subsection{Correctness of the validation protocol}
 $\Pi_c$ ensures consensus on CP blocks and facilitators,
which makes it the building block for reaching consensus on transactions.
In this section, we build on top of \Cref{theorem:consensus} to show that our validation protocol described in \Cref{sec:vd-protocol} has the agreement and validity properties of a TrustChain validation protocol defined in \Cref{def:validation}.

\begin{theorem}
\label{theorem:validation-agreement}
% (Agreement of the validation protocol)
$\Pi_v$ satisfies the agreement and validity properties of a TrustChain validation protocol (\Cref{def:validation}).
\end{theorem}
\begin{proof}
We proof the agreement property by contradiction.
From~\Cref{alg:get-validity},
we use the fact that $F_{u, i}$ and $t_{u, i}$ are valid.
Without loss of generality, suppose for some transaction with TX block $t$,
node $u$ decides \emph{valid} but node $v$ decides \emph{invalid}.
Then there exists a fragment $F' = \{ \dots, t', c'\}$ which $u$ received that contains a valid pair of $t$---$t'$.
There also exists a fragment $F'' = \{ \dots, t'', c''\}$ which $v$ received that does not contain or contains an invalid pair---$t''$.
In both cases, the $\textsf{get\_validity}(\cdot)$ function must have reached \Cref{line:valid-fragment}.
Due to \Cref{theorem:consensus}, we have $c' = c''$, otherwise the result would be \emph{unknown}.
Since $c' (= c'') = \langle \textsf{H}(t'), \dots \rangle$ we must have $\textsf{H}(t') = \textsf{H}(t'')$ and $t' \ne t''$ (because $t''$ is invalid).
In other words, the sender of $F''$ must be able to create some $t''$ that has the same digest as $t'$.
But this is only possible if the adversary can compute the inverse of $H(\cdot)$ with non-negligible probability.
Thus we have a contradiction and this completes the agreement proof.

@@

\subsection{Impossibility of liveness}
\Cref{theorem:validation-agreement} is a major result that allows significantly improved performance over traditional blockchain systems,
but it does not have all the properties of a typical Byzantine consensus problem.
Now we show a negative result, where the liveness property of \Cref{def:validation} cannot be attained in $\Pi_v$,
meaning that transactions made with adversaries cannot always be validated.

@@

\section{Performance analysis}
\label{sec:performance-analysis}
This section aims to analytically answer whether our system is horizontally scalable.
We begin by looking at the communication and time complexity of the consensus protocol $\Pi_c$,
and then the bandwidth requirement for a single transaction.
We build on top of those results to analyse the global throughput.

\subsection{Communication complexity of the consensus protocol}
\label{sec:cons-complexity}
$\Pi_c$ can be seen as three parts,
so the communication complexity is the sum of these parts.
The first part is when every node sends their CP block to all the facilitators, which is $O(Nn)$ since there are $N$ nodes and $n$ facilitators.
Or simply $O(N)$ if we consider $n$ as a constant.

@@

\section{Effect of a highly adversarial environment}
\label{sec:highly-adversarial}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=1.0\textwidth]{hypergeometric}
  \caption{Probability of selecting more than $\lfloor \frac{n-1}{3} \rfloor$ black balls for
  different value of $t$ with $N$ fixed at 2000.}
  \label{fig:hypergeometric}
\end{figure}

@@

The bound is not tight, but it is useful for picking parameters for a desired level of fault tolerance.
For some fixed $n$, since $0 \le \alpha \le \lfloor \frac{n-1}{3} \rfloor/n < (\lfloor  \frac{n - 1}{3} \rfloor + 1) / n$,
the probability is maximum when the squared term is minimum at $\alpha = \lfloor \frac{n-1}{3} \rfloor/n$.
The probability is minimum when the squared term is maximum at $\alpha = 0$.
Hence, if $n$ is known, then we can pick a $\alpha$ such that the probability becomes small.
On the other hand, if $\alpha$ is fixed, but small, we may increase $n$ to achieve the same.
To put this into perspective,
suppose $n = 1000$ and $\alpha = 1/5$, i.e. a fifth of the nodes are malicious.
Then the probability to draw more black balls than the threshold is only $2.6 \times 10^{-16}$.

### 2017-08-19 thesis/chapters/ch5-implementation.tex
Change code URL, etc.
Edit ch4 and 5

To build up the confidence of the horizontal scalability claim from our analytical results,
we must verify the same property experimentally.
Hence, we give an open source prototype implementation of our system.
Our implementation runs on a network with over 1000 node.
Each node sends transactions to each other autonomously, 
which allows us to evaluate our analytical claims.

We begin this chapter with a description of the implementation in \Cref{sec:implementation}.
Then, we move on to describing our experimental setup in \Cref{sec:experimental-setup} which includes a discussion on the system parameters and the physical infrastructure which we use.
Finally, \Cref{sec:comms-cost-experiment} and onwards follows the same structure as the Performance analysis (\Cref{sec:performance-analysis}),
where we analyse the performance experimentally rather than analytically and compare both results.

@@

The prototype implementation can be found on GitHub.
\begin{displayquote}
\url{https://github.com/kc1212/consensus-thesis-code}
\end{displayquote}
It implements the three protocols and the Extended TrustChain discussed in~\Cref{ch:model}.
We also implement two optimisations---privacy preserving validation protocol using compact blocks (\Cref{sec:compact})
and optimised validation protocol using cached agreed fragments (\Cref{sec:caching}).
It is written in the event driven paradigm, using the Python programming language\footnote{\url{https://www.python.org/}}.
We use the Twisted\footnote{\url{https://twistedmatrix.com/}} library for networking.

@@

The goal of the experiment is to run the three protocols---consensus protocol,
transaction protocol and validation protocol---simultaneously and analyse the communication cost, the consensus duration and the global throughput.
We investigate these properties under the following four parameters.
\begin{enumerate}
  \item The transaction rate $r_{\text{tx}}$ per node. This is not comparable to the others because it is fixed at 2 TX/s.
        Nevertheless, it is a good value because, as we show later,
        it hits a bottleneck in extreme cases which helps us understand the limitations of our design.
  \item The number of facilitators $n \in \{4, 8, \dots, 32\}$.
        The maximum $n$ is 32 because the limitation in liberasurecode mentioned in \Cref{sec:implementation},
        but our results give a good indication of how our system may function for larger numbers of $n$.
  \item The population size $N \in \{200, 300, \dots, 1200\}$.
        $N$ stops at 1200 is due to our physical setup, which we describe next.
  \item The two modes of transaction.
        The first mode is the \emph{fixed-neighbour} mode where nodes only transact with their immediate neighbour which is predetermined. 
        This minimises the data volume per validated transaction because agreed fragment can be cached.
        The second mode is in the other extreme which we call the \emph{random-neighbour} mode,
        where every transaction is with a random node out of the $N$ nodes in the system.
        It is unlikely that the agreed fragments can be reused.
\end{enumerate}

@@

Finally, since Bitcoin transactions are approximately 500 bytes~\cite{txsize},
we use a uniformly random transaction size sampled between 400 and 600 bytes.

@@

\begin{figure}[tb]
  \centering
  \makebox[\linewidth][c]{%
    \begin{subfigure}[b]{.7\textwidth}
      \includegraphics[width=\textwidth]{neighbour-fixed/round-duration-vs-population}
      \caption{Transactions are with fixed neighbours.}
      \label{fig:round-duration-fixed}
    \end{subfigure}%
    \begin{subfigure}[b]{0.7\textwidth}
      \includegraphics[width=\textwidth]{neighbour-random/round-duration-vs-population}
      \caption{Transactions are with random neighbours.}
      \label{fig:round-duration-random}
    \end{subfigure}%
  }
  \caption{Round duration does not increase linearly with the population size. 
  This is likely due to the additional hashing operation required for larger consensus result.}
  \label{fig:round-duration}
\end{figure}

\Cref{fig:round-comms} shows the relationship between the communication cost of the consensus protocol per round and the population size.
The most important aspect is that these results show a linear increase.
This reinforces our analytical result in~\Cref{sec:cons-complexity}.
Note that regardless of whether the transactions are performed with a random neighbour or with a fixed neighbour,
the magnitude of the communication cost is similar.
Both peak at about 100 MB.
This is expected because the consensus protocol is decoupled from the transaction protocol and the validation protocol.
Finally, the rate for which the communication cost increases is higher when the number of facilitators is higher.
This is also expected because the ACS algorithm has an $n^2$ term in its communication complexity.

@@

More interestingly, there is a large difference in communication cost between the two modes of transaction.
When transacting with only one neighbour, the communication cost is low because only one agreed fragment needs to be communicated for every round in order to validate all transactions of that round.
This is because agreed fragments are cached.
On the other hand,
if every node is transacting with a random node, then it is likely the case that one agreed fragment needs to be communicated for every transaction.
Hence the communication cost we see in~\Cref{fig:tx-comms-random} is much higher than in~\Cref{fig:tx-comms-fixed}.

@@

Note that the throughput decreases slightly as the number of facilitators increases.
This is due to the additional communication cost for running ACS with a high number of facilitators.
That is, if the network is congested then the nodes may not have enough bandwidth to send validation responses timely.

For~\Cref{fig:global-throughput-fixed},
the magnitude of our throughput may not be self-evident at first glance.
Recall that we fixed our $r_\text{tx}$ to 2, but how is it possible to have around 4800 transactions per second for 1200 nodes (which is 4 TX/s)?
This is due to the way validated transactions are calculated.
Transactions are between two parties, hence if every node makes two transactions per second,
every node also expects to receive two transactions per second.
Hence, for every node, the TX blocks are created at 4 per second.
Validation requests are sent at the same rate, which explains the magnitude.

>>>>

We evaluate our system experimentally to ultimately understand its scalability characteristics as well as limitations.
Concretely,
we give an open source prototype implementation that runs on a network with over 1000 node.
Nodes in our implementation make transactions autonomously with each other and then we collect the results for analysis.

This chapter begins with a description of the implementation in \Cref{sec:implementation}.
Then, we describe our experimental setup in \Cref{sec:experimental-setup},
which includes a discussion on the system parameters and the physical infrastructure which we use.
Finally, \Cref{sec:comms-cost-experiment} and onwards follows the same structure as \Cref{sec:performance-analysis},
where we analyse the performance experimentally rather than analytically and compare the results.

@@

The prototype implementation can be found on GitHub.
\begin{displayquote}
\url{https://github.com/kc1212/checo}
\end{displayquote}
It implements the three protocols and the Extended TrustChain discussed in~\Cref{ch:model}.
We also implement two optimisations---privacy preserving validation protocol using compact blocks (\Cref{sec:compact})
and optimised validation protocol using cached agreed fragments (\Cref{sec:caching}).
It is written in the event driven paradigm, using the Python programming language\footnote{\url{https://www.python.org/}}.
We use the Twisted\footnote{\url{https://twistedmatrix.com/}} library for networking.

@@

The goal of the experiment is to run the three protocols---consensus protocol,
transaction protocol and validation protocol---simultaneously and analyse the communication cost, the consensus duration and the global throughput.
We investigate these properties under the following parameters.
\begin{enumerate}
  \item The number of facilitators $n \in \{4, 8, \dots, 32\}$.
        The maximum $n$ is 32 because the limitation in liberasurecode mentioned in \Cref{sec:implementation},
        but our results give a good indication of how our system may function for larger numbers of $n$.
  \item The population size $N \in \{200, 300, \dots, 1200\}$.
        $N$ stops at 1200 is due to our physical setup, which we describe next.
  \item The two modes of transaction.
        The first mode is the \emph{fixed-neighbour} mode where nodes only transact with their immediate neighbour which is predetermined. 
        This minimises the data volume per validated transaction because agreed fragment can be cached.
        The second mode is in the other extreme which we call the \emph{random-neighbour} mode,
        where every transaction is with a random node out of the $N$ nodes in the system.
        It is unlikely that the agreed fragments can be reused.
\end{enumerate}

@@

Finally, the transaction rate $r_{\text{tx}}$ is fixed at 2 TX/s for every node.
This value, as we show later,
creates a bottleneck in extreme cases which helps us understand the limitations of our design.
Since Bitcoin transactions are approximately 500 bytes~\cite{txsize},
we use a uniformly random transaction size sampled between 400 and 600 bytes.

@@

\begin{figure}[tb]
  \centering
  \makebox[\linewidth][c]{%
    \begin{subfigure}[b]{.7\textwidth}
      \includegraphics[width=\textwidth]{neighbour-fixed/round-duration-vs-population}
      \caption{Transactions are with fixed neighbours.}
      \label{fig:round-duration-fixed}
    \end{subfigure}%
    \begin{subfigure}[b]{0.7\textwidth}
      \includegraphics[width=\textwidth]{neighbour-random/round-duration-vs-population}
      \caption{Transactions are with random neighbours.}
      \label{fig:round-duration-random}
    \end{subfigure}%
  }
  \caption{Round duration does not increase linearly with the population size,
  which is likely due to the additional hashing operation required for larger consensus result.}
  \label{fig:round-duration}
\end{figure}

\Cref{fig:round-comms} shows the relationship between the communication cost of the consensus protocol per round and the population size.
The most important aspect is that these results show a linear increase.
This reinforces our analytical result in~\Cref{sec:cons-complexity}.
Note that regardless of whether the transactions are performed with a random neighbour or with a fixed neighbour,
the magnitude of the communication cost is similar.
Both peak at about 100 MB,
which is expected because the consensus protocol is decoupled from the transaction protocol and the validation protocol.
Finally, the rate for which the communication cost increases is higher when the number of facilitators is higher,
which is also expected because the ACS algorithm has a polynomial term with respect to $n$ in its communication complexity.

@@

More interestingly, there is a large difference in communication cost between the two modes of transaction.
When transacting with only one neighbour, the communication cost is low because only one agreed fragment needs to be communicated for every round in order to validate all transactions of that round,
which is because agreed fragments are cached.
On the other hand,
if every node is transacting with a random node, then it is likely the case that one agreed fragment needs to be communicated for every transaction.
Hence the communication cost we see in~\Cref{fig:tx-comms-random} is much higher than in~\Cref{fig:tx-comms-fixed}.

@@

Note that the throughput decreases slightly as the number of facilitators increases,
which is due to the additional communication cost for running ACS with a high number of facilitators.
That is, if the network is congested then the nodes may not have enough bandwidth to send validation responses timely.

For~\Cref{fig:global-throughput-fixed},
the magnitude of our throughput may not be self-evident at first glance.
Recall that we fixed our $r_\text{tx}$ to 2, but how is it possible to have around 4800 transactions per second for 1200 nodes (4 TX/s)?
This result is due to the way validated transactions are calculated.
Transactions are between two parties, hence if every node makes two transactions per second,
every node also expects to receive two transactions per second.
Hence, for every node, the TX blocks are created at 4 per second.
Validation requests are sent at the same rate, which explains the magnitude.
